---
title: "Homozygote Freqs Figure for Perspectives"
output: 
  html_notebook:
    toc: true
    toc_float: true
---

Here I am going to make the figures for the perspectives piece that show the
expected vs observed homozygote frequencies.  I have put the 012 data that 
I will use into `./explorations/data_for_perspective_figure_1`

load libs
```{r}
library(genoscapeRtools)
library(stringr)
library(tidyverse)
```


## Read data in:

### Prince et al Chinook

I crunched this in eca-premature-migration-2017 and copied the results over.  It is 
genotypes called by ANGSD with a uniform prior.
```{r}
asnp <- read_tsv("data_for_perspective_figure_1/chinook-like-pca-012-transposed.txt.gz", col_names = FALSE)
asnp_names <- paste(asnp$X1,  asnp$X2, sep = "--")

angsd_mat <- as.matrix(asnp[,-c(1,2)])
rownames(angsd_mat) <- asnp_names
storage.mode(angsd_mat) <- "integer"
colnames(angsd_mat) <- NULL

angsd012 <- t(angsd_mat)[-191,]  # have to pull off the last row which was from reading an extra empty tab

chin_names <- scan("data_for_perspective_figure_1/chinook-bamlist-names.txt", what = "character")
rownames(angsd012) <- chin_names
```

Now, count up the number from each population:
```{r}
angsd_pops <- tibble(sample = rownames(angsd012)) %>%
  tidyr::separate(sample, into = c("species", "river", "mig_pattern", "N1", "N2"), remove = FALSE) %>%
  mutate(pop = str_c(river, "--", mig_pattern)) %>%
  select(sample, pop)

angsd_pops %>%
  count(pop)
```

Note that Tri-M has the largest sample so we will use that.
```{r}
use_em <- angsd_pops %>%
  filter(pop == "Tri--M")

trinM012 <- angsd012[use_em$sample,]

prince_gfc <- genoscapeRtools::geno_freq_calcs(trinM012, use_em) %>%
    filter(!is.nan(z_score))  # this chucks out monomorphic loci

prince012 <- trinM012

```


### hecht

```{r}
name <- "chinook_hecht"
d012 <- genoscapeRtools::read_012(prefix = file.path("data_for_perspective_figure_1", name), gz = TRUE)
pops <- readr::read_csv(paste0("data_for_perspective_figure_1/", name, "_pops.csv"))

bp_pops <- pops %>%
    filter(pop == "JOC")  # this is the pop with the largest sample 

bp_names <- bp_pops %>%
  .$sample

# this just takes data from the biggest pop samples
bp012 <- d012[bp_names,]

hecht012 <- bp012

hecht_gfc <- genoscapeRtools::geno_freq_calcs(bp012, bp_pops) %>%
    filter(!is.nan(z_score))  # this chucks out monomorphic loci
```

### wak_chinook

```{r}
name <- "wak_chinook"
d012 <- genoscapeRtools::read_012(prefix = file.path("data_for_perspective_figure_1", name), gz = TRUE)
pops <- readr::read_csv(paste0("data_for_perspective_figure_1/", name, "_pops.csv"))

bp_pops <- pops %>%
    filter(pop == "Kogrukluk")  # this is the pop with the largest sample 

bp_names <- bp_pops %>%
  .$sample

# this just takes data from the biggest pop samples
bp012 <- d012[bp_names,]

wak012 <- bp012

wak_gfc <- genoscapeRtools::geno_freq_calcs(bp012, bp_pops) %>%
    filter(!is.nan(z_score))  # this chucks out monomorphic loci
```

## combine those and plot them

```{r}
togeth <- list(
  prince = prince_gfc,
  wak = wak_gfc,
  hecht = hecht_gfc
) %>%
  bind_rows(.id = "dataset")

allhomo <- togeth %>%
  filter(geno != 1)
```

Plot:
```{r}
source("../../gbs-miscall-rates/R/geno-freq-boundaries.R")

tmp <- allhomo %>%
  mutate(dstr = recode(dataset,
                       wak = "(a) Korukluk,   n = 57,   L = 10,944,   HMR = 0.02",
                       hecht = "(b) Johnson Creek,   n = 68,  L = 13,764,   HMR = 0.21",
                       prince = "(c) Umpqua Mature,   n = 27,   L = 15,228,   HMR = 0.53" 
                       ))

g <- ggplot(tmp, aes(x = p_exp, y = p_obs)) + 
  facet_wrap(~dstr, ncol = 1) +
  geom_jitter(color = "blue", alpha = 0.05) +
  geom_polygon(data = geno_freq_boundaries() %>% filter(geno == 0), fill = NA, linetype = "dashed", colour = "black") +
  geom_abline(slope = 1, intercept = 0, colour = "black") +
  theme_bw() +
  xlab("Expected homozygote frequency") +
  ylab("Observed homozygote frequency")

ggsave(g, filename = "outputs/chinook_3_studies.pdf", width = 4, height = 9)
```



## Estimate the HMR

```{r}
# source the function that does it:
source("../../gbs-miscall-rates/R/estimate-m-etc.R")
```

### wak

```{r}
wak_hmr <- estimate_m(wak012, nreps = 500, m_init = 0.1)
```

```{r}
plot(wak_hmr$m)
mean(wak_hmr$m[-(1:100)])
```

### hecht

```{r}
hecht_hmr <- estimate_m(hecht012, nreps = 800, m_init = 0.1)
plot(hecht_hmr$m)
mean(hecht_hmr$m[-(1:200)])
```
 
 ### prince
 
```{r}
prince_hmr <- estimate_m(prince012, nreps = 1000, m_init = 0.1)
plot(prince_hmr$m)

mean(prince_hmr$m[-(1:400)])
```
